\Large \bf SIRIUS: A Revolutionary Approach to I/O and Storage for Exascale Scientific Computing

### 2016-03-11 papers/2016-HotStorage/paper.tex
added results and fixed error percentages in the example results table.

@@

\begin{table}[tbp]
\centering
\caption{Reading Options for Variable ``A''}
\label{tab:results}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\# & Size & Time & Time Err & Refactoring & Data Err\\
\hline
1 & $(\frac{1}{4})^3(\frac{3}{8})$A & 10s & $\pm$ 3s & stride, byte-split & 99\% \\
\hline
2 & $(\frac{3}{4})^3(\frac{3}{8})$A & 90s & $\pm$ 30s & stride, byte-split & 58\% \\
\hline
3 & $(\frac{1}{4})^3(\frac{5}{8})$A & 16s & $\pm$ 5s & stride, byte-split & 0.01\% \\
\hline
4 & $(\frac{3}{4})^3(\frac{5}{8})$A & 120s & $\pm$ 50s & stride, byte-split & 0.01\% \\
\hline
\hline
5 & $(\frac{1}{4})^3$A & 1200s & $\pm$ 30s & stride & 98\% \\
\hline
6 & $(\frac{3}{4})^3$A & 2400s & $\pm$ 90s & stride & 58\% \\
\hline
\hline
7 & $(\frac{3}{8})$A & 1350s & $\pm$ 120s & byte-split & 5\% \\
\hline
8 & $(\frac{5}{8})$A & 2250s & $\pm$ 120s & byte-split & 0.01\% \\
\hline
\hline
9 & A & 36s & $\pm$ 6s & wavelet & 1\% \\
\hline
\hline
10 & A & 3600s & $\pm$ 600s & none & 0\% \\
\hline
\end{tabular}
\end{table}

@@

This stack has an early prototype implementation intended to test concepts
rather than performance and scalability. It has focused on examining the
interaction of the different APIs for each layer to flesh out any detailed
requirements or concerns that may have been missed in the conceptualization of
this IO stack. To demonstrate the viability of the IO stack described in this
paper, we show some very early performance results from the untuned prototype.

@@

%\includegraphics[scale=0.07]{logos/doe_logo}
%\includegraphics[scale=0.30]{logos/snl_logo}
%\includegraphics[scale=0.35]{logos/nnsa_logo}
Sandia National Laboratories is a multi-program laboratory managed and operated
by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
Corporation, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.

>>>>

The byte splitting approach error rate is quite small. By splitting at byte
boundaries rather than converting to a lower precision 4-byte float, we
preserve the full exponent range and keep the first 11 bits of mantissa
precision. This leaves a $2^{-12}$ maximum error magnitude for each value--the
last 40 bits represent less precision together than the immediately preceeding
bit all by itself.

@@

\begin{table}[tbp]
\centering
\caption{Reading Options for Variable ``A''}
\label{tab:results}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\# & Size & Time & Time Err & Refactoring & Data Err\\
\hline
1 & $(\frac{1}{4})^3(\frac{3}{8})$A & 10s & $\pm$ 3s & stride, byte-split & 98\% \\
\hline
2 & $(\frac{3}{4})^3(\frac{3}{8})$A & 90s & $\pm$ 30s & stride, byte-split & 58\% \\
\hline
3 & $(\frac{1}{4})^3(\frac{5}{8})$A & 16s & $\pm$ 5s & stride, byte-split & n/a \\
\hline
4 & $(\frac{3}{4})^3(\frac{5}{8})$A & 120s & $\pm$ 50s & stride, byte-split & n/a \\
\hline
\hline
5 & $(\frac{1}{4})^3$A & 1200s & $\pm$ 30s & stride & 98\% \\
\hline
6 & $(\frac{3}{4})^3$A & 2400s & $\pm$ 90s & stride & 58\% \\
\hline
\hline
7 & $(\frac{3}{8})$A & 1350s & $\pm$ 120s & byte-split & 0.02\% \\
\hline
8 & $(\frac{5}{8})$A & 2250s & $\pm$ 120s & byte-split & n/a \\
\hline
\hline
9 & A & 36s & $\pm$ 6s & wavelet & 1\% \\
\hline
\hline
10 & A & 3600s & $\pm$ 600s & none & 0\% \\
\hline
\end{tabular}
\end{table}

@@

Our initial tests are on the Sith cluster at Oak Ridge. The system contains 40
compute nodes. Each compute node contains four 2.3 GHz 8 core AMD Opteron
processors, and 64 GB of memory. The system is configured with a 86 TB Lustre
file system for scratch space.

We use the XGC1~\cite{ku:2008:xgc1} fusion simulation as our application. We
run on 5 nodes with 160 processes total. XGC1 is configured to use 819,200,000
particles evenly spread across the processes.

\begin{table}[tbp]
\centering
\caption{Performance For 3/5 byte Split}
\label{tab:performance}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
Test & Time (s)\\
\hline
Write full data & 10.3 \\
\hline
Byte split & 2.5 \\
\hline
write 3-byte portion & 3.4 \\
\hline
write 5-byte portion & 9.3 \\
\hline
\hline
Read 3-byte portion & 10.3 \\
\hline
Reassemble after read & 27.9 \\
\hline
Read full data & 34.2 \\
\hline
\end{tabular}
\end{table}

Evaluating the error when using just the 3-byte portion compared to the full
data set yields an $L^2$ norm of 72028.2 and $L_{\inf}$ of 0.00109242.

Writing the 3-byte portion synchronously to fast storage and spooling the
5-byte portion out to slower storage asynchronously will preserve the full data
fidelity while improving both writing and reading time even when including the
disassemble/reassemble times. The introduced errors when using just the 3-byte
portion, as discussed above, is quite small.

@@

This project is funded by the following research grants from the Advanced
Scientific Research Office at the Department of Energy, under contracts SDAV:
ERKJ200/KJ0403000, RSVP: ERKJU60/KJ0402000, and Sirius: ERKJ311/KJ0402000.
Sandia National Laboratories is a multi-program laboratory managed and operated
by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
Corporation, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.
